---
title: "Regression Models Course Project"
author: "jacethedatascientist"
date: "August 13, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Regression Models Course Project
#### *Jace Galleon*

***

## I. Project Overview
In this project, we will work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome).  

We will examine the mtcars data set and explore how miles per gallon (MPG) is affected by different variables. In particularly, we will answer the following two questions: (1) Is an automatic or manual transmission better for MPG, and (2) Quantify the MPG difference between automatic and manual transmissions.

***

## II. Project Objective
Motor Trend is particularly interested in the following two questions:
1. “Is an automatic or manual transmission better for MPG”
2. "Quantify the MPG difference between automatic and manual transmissions"

## III. Project Analysis
### A. Data Preparation
We will use the **mtcars** dataset installed in the R Package and will take a sample.
```{r , cache = TRUE, echo = TRUE}
library(ggplot2)
data(mtcars)
head(mtcars)
```
There are other variables that are numerical in nature but are actually entered as *"types"* so we will convert them to factors.
```{r , cache = TRUE, echo = TRUE}
mtcars$cyl  <- factor(mtcars$cyl)
mtcars$vs   <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am   <- factor(mtcars$am,labels=c("Automatic","Manual"))
```

***

### B. Data Exploration
###### *Based on __Figure F-1__, there is an obvious difference between the impact of each Transmission Types to MPG. It can be seen that Automatic Transmission has a better impact compared to Manual.*
```{r , cache = TRUE, echo = TRUE}
aggregate(mpg~am, data = mtcars, mean)
```
On hypothesis, there is an average of 7.245 difference in the MPG between Manual and Automatic Transmission. But, to determine if there is a significant difference, we'll be using T-Test.

**Null** : *There is no significant difference between the mean of MPG for both Transmission Types.*  
**Alternative** : *There is a significant difference between the mean of MPG for both Transmission Types.*

```{r , cache = TRUE, echo = FALSE}
t.test(mtcars$mpg[mtcars$am == "Automatic"], mtcars$mpg[mtcars$am == "Manual"], paired = FALSE, var.equal = FALSE)
```

###### *Based on the P-value (0.001374), there is a significant difference between the mean of MPG for both Transmission Types.*


***

\newpage

### C. Data Analysis
```{r , cache = TRUE, echo = TRUE}
linmod <- lm(mpg ~ am, data = mtcars)
summary(linmod)
```
This summary shows us that the average MPG for cars with AT has an **average of 17.147** while for Manual is **7.245** higher.

If you look at the value of the _**R^2^**_, the value is **0.36**, which means that this accounts to **36% of the variation** in the MPG. To get a more accurate result, we'll be doing a multi-variate regression that will use the rest of the variables.

```{r , cache = TRUE, echo = TRUE}
multi_var <- lm(mpg~am + cyl + disp + hp + wt, data = mtcars)
anova(linmod,multi_var)
```
Based on this, we can infer that **cyl**, **disp**, **hp**, **wt** have a stronger correlation with mpg than **am**. We, then, build a new model using these variables and compare them to the initial model with the anova function to determine if there's a significant difference.

The P-value is **8.637e-08** and is almost 0, thus, we can infer that this new model is **better** than our initial (*linmod*).

Double-check the residuals for *non-normality* (see Figure F-2) and we can see that they are all *normally distributed*.


***

### D. Assumptions and Conclusions
```{r , cache = TRUE, echo = TRUE}
summary(multi_var)
```
The model accounts to **86.64%** of the variance and as a result, **cyl**, **disp**, **hp**, and **wt** did affect the correlation between mpg and am by **roughly 51%**. Thus, we can say that the difference between automatic and manual transmissions is 1.81 MPG with the manual being higher.

***

*This formally ends the Course Project. Thank You!*

***

\newpage

### E. Appendices
##### 1. Boxplot for the impact of the Transmission Type (am) to the MPG.
```{r , cache = TRUE, echo = FALSE}
boxplot(mpg ~ am, data = mtcars, col = (c("yellow","red")), ylab = "Miles per Gallon", xlab = "Transmission Type")
```

\newpage

##### 2. Plot of the latest model (multi_var) to show the distribution.
```{r , cache = TRUE, echo = FALSE}
par(mfrow = c(2,2))
plot(multi_var)
```
